---
title: "CRESS report for sequential extractions"
output: 
  github_document:
  html_preview: true
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

options(knitr.kable.NA = "")

```

```{r packages}

library(tidyverse)
library(googlesheets4)
library(whistledown)

theme_set(theme_bw())
options(scipen = 9999)

pal_soil_sequence = whistledown::whistledown_palette("kathani", 5)
pal_amd_sequence = whistledown::whistledown_palette("featherington", 7)
pal_tessier_sequence = whistledown::whistledown_palette("queen", 6)

```

```{r load_files}

icp_processed = read.csv("1-data/data_processed/ICP_processed.csv")
icp_digests = read.csv("1-data/data_processed/ICP_processed_digests.csv")
xrf_processed = read.csv("1-data/data_processed/XRF_processed.csv")

analytes_of_interest = c("Co", "Cr", "Cu", "Fe", "Li", "Ni", "Sc", "Sr", "Zn")
analytes_list = read_sheet("https://docs.google.com/spreadsheets/d/1KVmtSHtLs9ljyzNVzbpEQ-GWavy0_Bpi-gCd8MOw6uc/edit?gid=0#gid=0")

```


--- 

# Experimental design

<details>
<summary> Click to open </summary>

## Which analytes were measured?

Full list of analytes

```{r}

analytes_list %>%  
  knitr::kable()

```


Subset of analytes being discussed here

```{r}

analytes_list = read_sheet("https://docs.google.com/spreadsheets/d/1KVmtSHtLs9ljyzNVzbpEQ-GWavy0_Bpi-gCd8MOw6uc/edit?gid=0#gid=0") %>% 
  filter(analyte %in% analytes_of_interest)

analytes_list %>%  
  knitr::kable()

```


--- 


## Extractions performed:

- Single-point DTPA extraction
- "Soil sequence" sequential extraction: water --> HCl --> sodium dithionite --> sodium pyrophosphate
- "AMD sequence" (Acid Mine Drainage) sequential extraction
- "Tessier sequence" sequential extraction

Microwave digestions were performed to quantify total concentrations in the samples. 
The "residual fraction" was calculated by subtracting total extractable values (from the sequences) from the digest concentrations.



</details>



---

## ICP-MS results -- ug/g

```{r icp_files}

icp_summary = 
  icp_processed %>% 
  # keep only the analytes of interest
  filter(analyte %in% analytes_of_interest) %>% 
  # Sc and Sr in CRESS_Mud1 were suspect. so drop them
#  filter(!(CRESS_ID == "CRESS_Mud1" & (analyte == "Sc" | analyte == "Sr"))) %>% 
  filter(!(CRESS_ID == "CRESS_Mud1")) %>% 
  group_by(CRESS_ID, sample_name, extraction_sequence, fraction, group, analyte) %>% 
  dplyr::summarise(ugg = mean(ug_g),
                   ugg_se = sd(ug_g)/sqrt(n()),
                   mgg = ugg/1000,
                   percent = mean(percent)) %>% 
  mutate(across(where(is.numeric), round, 2))

```


### DTPA extraction (single point extraction)

This is the bioavailable fraction

```{r dtpa, fig.height=7, fig.width=10}

icp_summary %>% 
  filter(fraction == "A: DTPA-available") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

- Sc and Cr were not measured for soil extracts (because the necessary standard was not available)
- Li was below detection for all


### Sequential extraction (soil sequence)

```{r soil_sequence, fig.height=7, fig.width=10}

icp_summary %>% 
  filter(extraction_sequence == "Soil Sequence") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  scale_fill_manual(values = pal_soil_sequence)+ 
  scale_y_continuous(labels = scales::comma)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

- Sc and Cr were not measured for soil extracts (because the necessary standard was not available)


### Sequential extraction (AMD sequence)

```{r amd_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "AMD Sequence") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  scale_fill_manual(values = pal_amd_sequence)+ 
  scale_y_continuous(labels = scales::comma)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (Tessier sequence)

```{r tessier_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "Tessier Sequence") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  scale_fill_manual(values = pal_tessier_sequence)+ 
  scale_y_continuous(labels = scales::comma)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

---


---

## ICP-MS results -- percentage

### Sequential extraction (soil sequence)

```{r percent_soil_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "Soil Sequence") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
  scale_fill_manual(values = pal_soil_sequence)+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (AMD sequence)

```{r percent_amd_sequence, fig.height=7, fig.width=10}

icp_summary %>% 
  filter(extraction_sequence == "AMD Sequence") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
  scale_fill_manual(values = pal_amd_sequence)+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (Tessier sequence)

```{r percent_tessier_sequence, fig.height=7, fig.width=10}

icp_summary %>% 
  filter(extraction_sequence == "Tessier Sequence") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
  scale_fill_manual(values = pal_tessier_sequence)+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

---

# XRF analysis

```{r fig.height=10, fig.width=14}

xrf_processed %>% 
  ggplot(aes(x = CRESS_ID, y = XRF_ugg))+
  geom_point()+
  facet_wrap(~analyte, ncol = 4, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## Comparing XRF data with digests

```{r fig.height=10, fig.width=14}

xrf_digest_compar = 
  xrf_processed %>% 
  full_join(icp_digests %>% dplyr::select(CRESS_ID, analyte, ug_g)) %>% 
  rename(digest_ugg = ug_g) %>% 
  arrange(analyte) %>% 
#  pivot_longer(cols = -c(CRESS_ID, analyte)) %>% 
  force()

xrf_digest_compar %>% 
  drop_na() %>% 
  ggplot(aes(x = XRF_ugg, y = digest_ugg, color = CRESS_ID))+
  geom_point(size = 2)+
  facet_wrap(~analyte, scales = "free")

```


---

```{r, eval=FALSE}
library(ggh4x)

icp_summary %>% 
  filter(extraction_sequence != "DTPA") %>% 
  ggplot(aes(x = extraction_sequence, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "analyte concentration (μg/g soil)")+
  facet_nested_wrap(~analyte + CRESS_ID, 
                    scales = "free_y"
                    )+
#        scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))


icp_summary %>% 
  filter(extraction_sequence != "DTPA") %>% 
  ggplot(aes(x = extraction_sequence, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "analyte concentration (μg/g soil)")+
  facet_grid(analyte ~ CRESS_ID, scales = "free_y")+
  scale_y_continuous(labels = scales::comma)+
  scale_fill_manual(values = whistledown_palette("queen", 15))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



# Li: Mud1
# Sc, Sr: Mud 1

```


---

## Session Info 

<details>
  <summary> Session Info </summary>

Date run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>