---
title: "CRESS report for sequential extractions"
output: 
  github_document:
  html_preview: true
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

```{r}

library(tidyverse)
theme_set(theme_bw())
options(scipen = 9999)

```


Extractions performed:

- Single-point DTPA extraction
- "Soil sequence" sequential extraction: water --> HCl --> sodium dithionite --> sodium pyrophosphate

---

## ICP-MS results

```{r}

icp_processed = read.csv("1-data/data_processed/ICP_processed.csv")

icp_summary = 
  icp_processed %>% 
  group_by(sample_name, extraction_sequence, extract, group, analyte) %>% 
  dplyr::summarise(ugg = mean(ug_g),
                   ugg_se = sd(ug_g)/sqrt(n()),
                   mgg = ugg/1000) %>% 
  mutate(across(where(is.numeric), round, 2))

```


### DTPA extraction (single point extraction)

```{r, fig.height=7, fig.width=10}

icp_summary %>% 
  filter(extract == "A: DTPA" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extract == "A: DTPA" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extract == "A: DTPA" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

### Sequential extraction (soil sequence)

```{r, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "Soil Sequence") %>% 
  filter(extract != "A: DTPA" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extract != "A: DTPA" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extract != "A: DTPA" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (AMD sequence)

```{r, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "AMD Sequence") %>% 
  filter(extract != "A: DTPA" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "AMD Sequence") %>% 
filter(extract != "A: DTPA" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extraction_sequence == "AMD Sequence") %>% 
  filter(extract != "A: DTPA" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (Tessier sequence)

```{r, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "Tessier Sequence") %>% 
  filter(extract != "A: DTPA" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "Tessier Sequence") %>% 
filter(extract != "A: DTPA" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extraction_sequence == "Tessier Sequence") %>% 
  filter(extract != "A: DTPA" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

---

## All reps

<details>
<summary> See plots for all reps here</summary>

### DTPA extraction (single point extraction)

```{r, fig.height=9, fig.width=15}

icp_processed %>% 
  filter(extract == "A: DTPA" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_ID, y = ug_g, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "cations")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_processed %>% 
  filter(extract == "A: DTPA" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_ID, y = ug_g, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_processed %>% 
  filter(extract == "A: DTPA" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_ID, y = ug_g, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "REEs")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

### Sequential extraction (soil sequence)

```{r, fig.height=9, fig.width=15}
icp_processed %>% 
  filter(extract != "A: DTPA" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_ID, y = ug_g, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "cations")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_processed %>% 
  filter(extract != "A: DTPA" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_ID, y = ug_g, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_processed %>% 
  filter(extract != "A: DTPA" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_ID, y = ug_g, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "REEs")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


--- 

</details>



--- 

<details>
<summary> See raw extract concentrations here </summary>

## ICP-MS results --- ug/L

```{r}

icp_processed = read.csv("1-data/data_processed/ICP_processed.csv")

icp_summary = 
  icp_processed %>% 
  group_by(sample_name, extract, group, analyte) %>% 
  dplyr::summarise(ppb = mean(ppb_blank_corr),
                   ppb_se = sd(ppb_blank_corr)/sqrt(n()),
                   ppm = ppb/1000) %>% 
  mutate(across(where(is.numeric), round, 2))

```


### DTPA extraction (single point extraction)

```{r, fig.height=7, fig.width=10}

icp_summary %>% 
  filter(extract == "A: DTPA" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_name, y = ppb/1000, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "cations")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extract == "A: DTPA" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_name, y = ppb/1000, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extract == "A: DTPA" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_name, y = ppb, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "REEs")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

### Sequential extraction (soil sequence)

```{r, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extract != "A: DTPA" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_name, y = ppb/1000, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "cations")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extract != "A: DTPA" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_name, y = ppb/1000, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extract != "A: DTPA" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_name, y = ppb, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "REEs")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

---

<details>
<summary> See plots for all reps here</summary>

### DTPA extraction (single point extraction)

```{r, fig.height=9, fig.width=15}

icp_processed %>% 
  filter(extract == "A: DTPA" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_ID, y = ppb_blank_corr/1000, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "cations")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_processed %>% 
  filter(extract == "A: DTPA" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_ID, y = ppb_blank_corr/1000, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_processed %>% 
  filter(extract == "A: DTPA" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_ID, y = ppb_blank_corr, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "REEs")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

### Sequential extraction (soil sequence)

```{r, fig.height=9, fig.width=15}
icp_processed %>% 
  filter(extract != "A: DTPA" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_ID, y = ppb_blank_corr/1000, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "cations")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_processed %>% 
  filter(extract != "A: DTPA" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_ID, y = ppb_blank_corr/1000, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_processed %>% 
  filter(extract != "A: DTPA" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_ID, y = ppb_blank_corr, fill = extract))+
  geom_bar(stat = "identity")+
  labs(title = "REEs")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


--- 

</details>


</details>




---

<details>
<summary> Session Info </summary>

Date run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>