---
title: "CRESS report for sequential extractions"
output: 
  github_document:
  html_preview: true
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

options(knitr.kable.NA = "")

```

```{r packages}

library(tidyverse)
library(googlesheets4)
library(whistledown)

theme_set(theme_bw())
options(scipen = 9999)

pal_soil_sequence = whistledown::whistledown_palette("kathani", 5)
pal_amd_sequence = whistledown::whistledown_palette("featherington", 7)
pal_tessier_sequence = whistledown::whistledown_palette("queen", 6)

```

```{r load_files}

icp_processed = read.csv("1-data/data_processed/ICP_processed.csv")
icp_digests = read.csv("1-data/data_processed/ICP_processed_digests.csv")
xrf_processed = read.csv("1-data/data_processed/XRF_processed.csv")

analytes_list = read_sheet("https://docs.google.com/spreadsheets/d/1KVmtSHtLs9ljyzNVzbpEQ-GWavy0_Bpi-gCd8MOw6uc/edit?gid=0#gid=0")

```


--- 

# Experimental design

<details>
<summary> Click to open </summary>

## Which analytes were measured?

```{r}

analytes_list %>%  
  knitr::kable()

```

--- 


## Extractions performed:

- Single-point DTPA extraction
- "Soil sequence" sequential extraction: water --> HCl --> sodium dithionite --> sodium pyrophosphate
- "AMD sequence" (Acid Mine Drainage) sequential extraction
- "Tessier sequence" sequential extraction

Microwave digestions were performed to quantify total concentrations in the samples. 
The "residual fraction" was calculated by subtracting total extractable values (from the sequences) from the digest concentrations.



</details>



---

## ICP-MS results -- ug/g

```{r icp_files}

icp_summary = 
  icp_processed %>% 
  group_by(CRESS_ID, sample_name, extraction_sequence, fraction, group, analyte) %>% 
  dplyr::summarise(ugg = mean(ug_g),
                   ugg_se = sd(ug_g)/sqrt(n()),
                   mgg = ugg/1000,
                   percent = mean(percent)) %>% 
  mutate(across(where(is.numeric), round, 2))

```


### DTPA extraction (single point extraction)

```{r dtpa, fig.height=7, fig.width=10}

icp_summary %>% 
  filter(fraction == "A: DTPA-available" & 
           group == "cations") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(fraction == "A: DTPA-available" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(fraction == "A: DTPA-available" & 
           group == "REEs") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

### Sequential extraction (soil sequence)

```{r soil_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "Soil Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "cations") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
    scale_fill_manual(values = pal_soil_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "Soil Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
      scale_fill_manual(values = pal_soil_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "Soil Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "REEs") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
      scale_fill_manual(values = pal_soil_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (AMD sequence)

```{r amd_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "AMD Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "cations") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
      scale_fill_manual(values = pal_amd_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "AMD Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
        scale_fill_manual(values = pal_amd_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extraction_sequence == "AMD Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "REEs") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
        scale_fill_manual(values = pal_amd_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (Tessier sequence)

```{r tessier_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "Tessier Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "cations") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
        scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "Tessier Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
          scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extraction_sequence == "Tessier Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "REEs") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
          scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

---


---

## ICP-MS results -- percentage

### Sequential extraction (soil sequence)

```{r percent_soil_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "Soil Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "cations") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
    scale_fill_manual(values = pal_soil_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "Soil Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
      scale_fill_manual(values = pal_soil_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "Soil Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "REEs") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
      scale_fill_manual(values = pal_soil_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (AMD sequence)

```{r percent_amd_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "AMD Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "cations") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
      scale_fill_manual(values = pal_amd_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "AMD Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
        scale_fill_manual(values = pal_amd_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extraction_sequence == "AMD Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "REEs") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
        scale_fill_manual(values = pal_amd_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (Tessier sequence)

```{r percent_tessier_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "Tessier Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "cations") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
        scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "Tessier Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
          scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extraction_sequence == "Tessier Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "REEs") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
          scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

---

# XRF analysis

```{r fig.height=10, fig.width=14}

xrf_digest_compar = 
  xrf_processed %>% 
  full_join(icp_digests %>% dplyr::select(CRESS_ID, analyte, ug_g)) %>% 
  rename(digest_ugg = ug_g) %>% 
  arrange(analyte) %>% 
#  pivot_longer(cols = -c(CRESS_ID, analyte)) %>% 
  force()

xrf_digest_compar %>% 
  drop_na() %>% 
  ggplot(aes(x = XRF_ugg, y = digest_ugg, color = CRESS_ID))+
  geom_point()+
  facet_wrap(~analyte, scales = "free")

```



---

```{r}
library(ggh4x)

icp_summary %>% 
 filter(extraction_sequence != "DTPA") %>% 
  filter(group == "CMMs and metals") %>% 
  ggplot(aes(x = extraction_sequence, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
              y = "analyte concentration (μg/g soil)")+
  facet_nested_wrap(~analyte + CRESS_ID, 
                    scales = "free_y"
                    )+
#        scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))


icp_summary %>% 
 filter(extraction_sequence != "DTPA") %>% 
  filter(group == "REEs") %>% 
  ggplot(aes(x = extraction_sequence, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
              y = "analyte concentration (μg/g soil)")+
  facet_nested_wrap(~analyte + CRESS_ID, 
                    scales = "free_y"
                    )+
#        scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

```{r}

icp_qc = 
  icp_processed %>%
  group_by(CRESS_ID, sample_name, extraction_sequence, group, analyte) %>% 
  dplyr::summarise(total_ugg = mean(total_ugg)) %>% 
  pivot_wider(names_from = "extraction_sequence", values_from = "total_ugg") %>% 
  mutate_if(where(is.numeric), round, 2) %>% 
  mutate(ALL_EQUAL = case_when((`AMD Sequence` == `Soil Sequence`) &(`Soil Sequence` == `Tessier Sequence`) ~ "Y", .default = "N"))
  
  



```

