---
title: "CRESS report for sequential extractions"
output: 
  github_document:
  html_preview: true
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
```

```{r packages}

library(tidyverse)
theme_set(theme_bw())
options(scipen = 9999)

pal_soil_sequence = whistledown::whistledown_palette("kathani", 5)
pal_amd_sequence = whistledown::whistledown_palette("featherington", 7)
pal_tessier_sequence = whistledown::whistledown_palette("queen", 6)

```


Extractions performed:

- Single-point DTPA extraction
- "Soil sequence" sequential extraction: water --> HCl --> sodium dithionite --> sodium pyrophosphate

---

## ICP-MS results -- ug/g

```{r icp_files}

icp_processed = read.csv("1-data/data_processed/ICP_processed.csv")

icp_summary = 
  icp_processed %>% 
  group_by(sample_name, extraction_sequence, fraction, group, analyte) %>% 
  dplyr::summarise(ugg = mean(ug_g),
                   ugg_se = sd(ug_g)/sqrt(n()),
                   mgg = ugg/1000,
                   percent = mean(percent)) %>% 
  mutate(across(where(is.numeric), round, 2))

```


### DTPA extraction (single point extraction)

```{r dtpa, fig.height=7, fig.width=10}

icp_summary %>% 
  filter(fraction == "A: DTPA-available" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(fraction == "A: DTPA-available" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(fraction == "A: DTPA-available" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

### Sequential extraction (soil sequence)

```{r soil_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "Soil Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
    scale_fill_manual(values = pal_soil_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "Soil Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
      scale_fill_manual(values = pal_soil_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "Soil Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
      scale_fill_manual(values = pal_soil_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (AMD sequence)

```{r amd_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "AMD Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
      scale_fill_manual(values = pal_amd_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "AMD Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
        scale_fill_manual(values = pal_amd_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extraction_sequence == "AMD Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
        scale_fill_manual(values = pal_amd_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (Tessier sequence)

```{r tessier_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "Tessier Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
              y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
        scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "Tessier Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
          scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extraction_sequence == "Tessier Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_name, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "analyte concentration (μg/g soil)")+
  facet_wrap(~analyte, scales = "free_y")+
          scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

---


---

## ICP-MS results -- percentage

### Sequential extraction (soil sequence)

```{r percent_soil_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "Soil Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_name, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
    scale_fill_manual(values = pal_soil_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "Soil Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_name, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
      scale_fill_manual(values = pal_soil_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "Soil Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_name, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
      scale_fill_manual(values = pal_soil_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (AMD sequence)

```{r percent_amd_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "AMD Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_name, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
      scale_fill_manual(values = pal_amd_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "AMD Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_name, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
        scale_fill_manual(values = pal_amd_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extraction_sequence == "AMD Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_name, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
        scale_fill_manual(values = pal_amd_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (Tessier sequence)

```{r percent_tessier_sequence, fig.height=7, fig.width=10}
icp_summary %>% 
  filter(extraction_sequence == "Tessier Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "cations") %>% 
  ggplot(aes(x = sample_name, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "cations",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
        scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
    filter(extraction_sequence == "Tessier Sequence") %>% 
filter(fraction != "A: DTPA-available" & 
           group == "CMMs and metals") %>% 
  ggplot(aes(x = sample_name, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "CMMs and metals",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
          scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

icp_summary %>% 
  filter(extraction_sequence == "Tessier Sequence") %>% 
  filter(fraction != "A: DTPA-available" & 
           group == "REEs") %>% 
  ggplot(aes(x = sample_name, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(title = "REEs",
       y = "percentage")+
  facet_wrap(~analyte, scales = "free_y")+
          scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

---
