---
title: "CRESS report for sequential extractions"
output: 
  github_document:
  html_preview: true
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

options(knitr.kable.NA = "")

```

```{r packages}

library(tidyverse)
library(googlesheets4)
library(whistledown)

theme_set(theme_bw())
options(scipen = 9999)

pal_soil_sequence = whistledown::whistledown_palette("kathani", 5)
pal_amd_sequence = whistledown::whistledown_palette("featherington", 7)
pal_tessier_sequence = whistledown::whistledown_palette("queen", 6)

theme_kp <- function() {  # this for all the elements common across plots
  theme_bw() %+replace%
    theme(legend.position = "right",
          legend.key=element_blank(),
          legend.title = element_text(size = 14, hjust = 0),
          legend.text = element_text(size = 12),
          legend.key.size = unit(1.5, 'lines'),
          legend.background = element_rect(colour = NA),
          panel.border = element_rect(color="black",linewidth=1, fill = NA),
          
          plot.title = element_text(hjust = 0, size = 14),
          axis.text = element_text(size = 14, color = "black"),
          axis.title = element_text(size = 14, face = "bold", color = "black"),
          
          # formatting for facets
          panel.background = element_blank(),
          strip.background = element_rect(colour="white", fill="white"), #facet formatting
          panel.spacing.x = unit(1, "lines"), #facet spacing for x axis
          panel.spacing.y = unit(1.5, "lines"), #facet spacing for x axis
          strip.text.x = element_text(size=14, face="bold"), #facet labels
          strip.text.y = element_text(size=14, face="bold", angle = 270) #facet labels
    )
}
theme_set(theme_kp())


```

```{r load_files}

icp_processed = read.csv("1-data/data_processed/ICP_processed.csv")
icp_digests = read.csv("1-data/data_processed/ICP_processed_digests.csv")
xrf_processed = read.csv("1-data/data_processed/XRF_processed.csv")

analytes_of_interest = c("Al", "Co", "Cu", "Fe", "Li", "Mn", "Ni", "Sr", "Zn")
analytes_list = read_sheet("https://docs.google.com/spreadsheets/d/1KVmtSHtLs9ljyzNVzbpEQ-GWavy0_Bpi-gCd8MOw6uc/edit?gid=0#gid=0")

```


--- 

# Experimental design

<details>
<summary> Click to open </summary>

## Which analytes were measured?

Full list of analytes

```{r}

analytes_list %>%  
  knitr::kable()

```


Subset of analytes being discussed here

```{r}

analytes_list = read_sheet("https://docs.google.com/spreadsheets/d/1KVmtSHtLs9ljyzNVzbpEQ-GWavy0_Bpi-gCd8MOw6uc/edit?gid=0#gid=0") %>% 
  filter(analyte %in% analytes_of_interest)

analytes_list %>%  
  dplyr::select(analyte, XRF, ICP) %>% 
  knitr::kable()

```


--- 


## Extractions performed:

- Single-point DTPA extraction
- "Soil sequence" sequential extraction: water --> HCl --> sodium dithionite --> sodium pyrophosphate
- "AMD sequence" (Acid Mine Drainage) sequential extraction
- "Tessier sequence" sequential extraction

Microwave digestions were performed to quantify total concentrations in the samples. 
The "residual fraction" was calculated by subtracting total extractable values (from the sequences) from the digest concentrations.



</details>



---

## ICP-MS results -- ug/g

```{r icp_files}

icp_summary = 
  icp_processed %>% 
  # keep only the analytes of interest
  filter(analyte %in% analytes_of_interest) %>% 
  # Sc and Sr in CRESS_Mud1 were suspect. so drop them
  filter(!(CRESS_ID == "CRESS_Mud1")) %>% 
  filter(!(CRESS_ID == "CRESS_Soil1")) %>% 
  # keep only soil samples
  filter(grepl("Soil", CRESS_ID)) %>% 
  mutate(CRESS_ID = str_remove(CRESS_ID, "CRESS_")) %>% 
  group_by(CRESS_ID, sample_name, extraction_sequence, fraction, group, analyte) %>% 
  dplyr::summarise(ugg = mean(ug_g),
                   ugg_se = sd(ug_g)/sqrt(n()),
                   mgg = ugg/1000,
                   percent = mean(percent)) %>% 
  mutate(across(where(is.numeric), round, 2))

```


### DTPA extraction (single point extraction)

This is the bioavailable fraction

```{r dtpa, fig.height=9, fig.width=12}

icp_summary %>% 
  filter(fraction == "A: DTPA-available") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "analyte concentration (μg/g soil)",
       x = "")+
  facet_wrap(~analyte, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

- Sc and Cr were not measured for soil extracts (because the necessary standard was not available)
- Li was below detection for all


### Sequential extraction (soil sequence)

```{r soil_sequence, fig.height=9, fig.width=12}

icp_summary %>% 
  filter(extraction_sequence == "Soil Sequence") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "analyte concentration (μg/g soil)",
       x = "")+
  facet_wrap(~analyte, scales = "free_y")+
  scale_fill_manual(values = pal_soil_sequence)+ 
  scale_y_continuous(labels = scales::comma)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

- Sc and Cr were not measured for soil extracts (because the necessary standard was not available)


### Sequential extraction (AMD sequence)

```{r amd_sequence, fig.height=9, fig.width=12}
icp_summary %>% 
  filter(extraction_sequence == "AMD Sequence") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "analyte concentration (μg/g soil)",
       x = "")+
  facet_wrap(~analyte, scales = "free_y")+
  scale_fill_manual(values = pal_amd_sequence)+ 
  scale_y_continuous(labels = scales::comma)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (Tessier sequence)

```{r tessier_sequence, fig.height=9, fig.width=12}
icp_summary %>% 
  filter(extraction_sequence == "Tessier Sequence") %>% 
  ggplot(aes(x = CRESS_ID, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "analyte concentration (μg/g soil)",
       x = "")+
  facet_wrap(~analyte, scales = "free_y")+
  scale_fill_manual(values = pal_tessier_sequence)+ 
  scale_y_continuous(labels = scales::comma)+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

---


---

## ICP-MS results -- percentage

### Sequential extraction (soil sequence)

```{r percent_soil_sequence, fig.height=9, fig.width=12}
icp_summary %>% 
  filter(extraction_sequence == "Soil Sequence") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "percentage",
       x = "")+
  facet_wrap(~analyte, scales = "free_y")+
  scale_fill_manual(values = pal_soil_sequence)+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (AMD sequence)

```{r percent_amd_sequence, fig.height=9, fig.width=12}

icp_summary %>% 
  filter(extraction_sequence == "AMD Sequence") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "percentage",
       x = "")+
  facet_wrap(~analyte, scales = "free_y")+
  scale_fill_manual(values = pal_amd_sequence)+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


### Sequential extraction (Tessier sequence)

```{r percent_tessier_sequence, fig.height=9, fig.width=12}

icp_summary %>% 
  filter(extraction_sequence == "Tessier Sequence") %>% 
  ggplot(aes(x = CRESS_ID, y = percent, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "percentage",
       x = "")+
  facet_wrap(~analyte, scales = "free_y")+
  scale_fill_manual(values = pal_tessier_sequence)+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

---

# XRF analysis

```{r fig.height=10, fig.width=14}

xrf_processed %>% 
  ggplot(aes(x = CRESS_ID, y = XRF_ugg))+
  geom_point()+
  facet_wrap(~analyte, ncol = 4, scales = "free_y")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## Comparing XRF data with digests

```{r fig.height=8, fig.width=10}

xrf_digest_compar = 
  xrf_processed %>% 
  full_join(icp_digests %>% dplyr::select(CRESS_ID, analyte, ug_g)) %>% 
  rename(digest_ugg = ug_g) %>% 
  arrange(analyte) %>% 
  filter(grepl("Soil", CRESS_ID)) %>% 
#  pivot_longer(cols = -c(CRESS_ID, analyte)) %>% 
  force()

xrf_digest_compar %>% 
  drop_na() %>% 
  filter(analyte %in% analytes_of_interest) %>% 
  ggplot(aes(x = XRF_ugg, y = digest_ugg, color = CRESS_ID))+
  geom_point(size = 4, show.legend = F)+
  geom_abline(slope = 1, linetype = "dashed", color = "grey50")+
  expand_limits(y = 0)+
  expand_limits(x = 0)+
  labs(x = "XRF results (μg/g)",
       y = "Digestion results (μg/g)")+
  facet_wrap(~analyte, scales = "free")

```


---

```{r, eval=FALSE}
library(ggh4x)

icp_summary %>% 
  filter(extraction_sequence != "DTPA") %>% 
  ggplot(aes(x = extraction_sequence, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "analyte concentration (μg/g soil)")+
  facet_nested_wrap(~analyte + CRESS_ID, 
                    scales = "free_y"
                    )+
#        scale_fill_manual(values = pal_tessier_sequence)+ 
theme(axis.text.x = element_text(angle = 45, hjust = 1))


icp_summary %>% 
  filter(extraction_sequence != "DTPA") %>% 
  ggplot(aes(x = extraction_sequence, y = ugg, fill = fraction))+
  geom_bar(stat = "identity")+
  labs(y = "analyte concentration (μg/g soil)")+
  facet_grid(analyte ~ CRESS_ID, scales = "free_y")+
  scale_y_continuous(labels = scales::comma)+
  scale_fill_manual(values = whistledown_palette("queen", 15))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



# Li: Mud1
# Sc, Sr: Mud 1

```


---

## Session Info 

<details>
  <summary> Session Info </summary>

Date run: `r Sys.Date()`

```{r}
sessionInfo()
```

</details>